---
title: "Omicron Covid Variant"
author: "Shubham"
date: "27/01/2022"
output:
  html_document: default
  pdf_document: default
---
# Loading packages
```{r}
library(tidyverse)
library(lubridate)
library(skimr)    # for statistical summary
library(janitor) # for examining and cleaning dirty data.
library(scales)
library(dplyr)
data<-read.csv("C:/Users/dell/OneDrive/Documents/covid-variants.csv")
```
*** Summary of Covid Data***
```{r}
glimpse(data)
summary(data)
str(data)
head(data)
```


```{r}
skim_without_charts(data)
data->covid_df
```

# Rename column
```{r}

covid_df <- covid_df %>%
  select(location, date, variant, num_sequences, perc_sequences, num_sequences_total) %>%
  rename(num_seq_processed = num_sequences, perc_seq_processed = perc_sequences)

colnames(covid_df)
```

# Unique Variants
```{r}
Uniq_variant = as.list(unique(covid_df$variant))
Uniq_variant

# group by variant
variant_df <- covid_df %>%
  group_by(variant) %>%
  summarize(total_seq_processed = sum(num_seq_processed))

head(variant_df)
```

# variant plot 
```{r}
options(repr.plot.width = 10, repr.plot.height = 15)
ggplot(data = variant_df, mapping = aes(x = total_seq_processed, y = variant, fill = variant)) + 
  geom_col(position = "dodge", color = "Black") + 
  geom_text(aes(label = total_seq_processed),hjust = -0.05, nudge_x = 0.5, size = 4, angle = 0) +
  labs(title = "COVID Variants vs Total number of Sequences Processed", x = "Total Sequences Processed", y = "Covid Variants", fill = "Covid Variants") +
  scale_x_continuous(labels = scales::comma, limit = c(0, 4500000)) + 
  theme_minimal() +
  theme(axis.text = element_text(size = 12)) +
  theme(axis.title = element_text(size = 20), title = element_text(size = 20))
```

# Total sequences processed <100,000
```{r}
options(repr.plot.width = 15, repr.plot.height = 15)
variant_df %>%
  filter(total_seq_processed < 100000) %>%
  ggplot(aes(x = variant, y = total_seq_processed, fill = variant)) + 
  geom_col(position = "dodge", color = "Black") + 
  geom_text(aes(label = total_seq_processed),vjust = -0.2, nudge_y = 0.5, size = 4, angle = 0) +
  labs(title = "COVID Variants vs Total number of Sequences Processed < 100,000", x = "Covid Variants", y = "Total Sequences Processed", fill = "Covid Variants") +
  scale_y_continuous(labels = scales::comma, limits = c(0, NA)) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle=90)) +
  theme(axis.text = element_text(size = 12)) +
  theme(axis.title = element_text(size = 20), title = element_text(size = 20))
```


# Total sequences processed >100,000
```{r}
options(repr.plot.width = 15, repr.plot.height = 15)
variant_df %>%
  filter(total_seq_processed > 100000) %>%
  ggplot(aes(x = variant, y = total_seq_processed, fill = variant)) + 
  geom_col(position = "dodge", color = "Black") + 
  geom_text(aes(label = total_seq_processed), vjust = -0.2, nudge_y = 0.5, size = 4, angle = 0) +
  labs(title = "COVID Variants vs Total number of Sequences Processed > 100,000", x = "Covid Variants", y = "Total Sequences Processed", fill = "Covid Variants") +
  scale_y_continuous(labels = scales::comma, limits = c(NA, NA)) + 
  theme_minimal() +
  theme(axis.text = element_text(size = 12)) +
  theme(axis.title = element_text(size = 20), title = element_text(size = 20))
```

# group by variant and location
```{r}
loc_variant_df <- covid_df %>%
  group_by(location, variant) %>%
  summarize(total_seq_processed = sum(num_seq_processed)) %>%
  filter(total_seq_processed >0)
head(loc_variant_df)

# Variant, location plot
options(repr.plot.width = 15, repr.plot.height = 30)
loc_variant_df %>%
  group_by(location, variant) %>%
  ggplot(mapping = aes(x = total_seq_processed, y = location, fill = variant)) + 
  geom_bar(stat = "identity") + 
  labs(title = "Countrywise COVID Variants vs Total number of Sequences Processed", x = "Total Sequences Processed", y = "Location", fill = "Covid Variants") +
  scale_x_continuous(labels = scales::comma) +
  theme(axis.text = element_text(size = 12)) +
  theme(axis.title = element_text(size = 20), title = element_text(size = 20))
```


# Group by location in desc. order
```{r}
loc_variant_df2 <- covid_df %>%
  group_by(location) %>%
  summarize(total_seq_processed = sum(num_seq_processed)) %>%
  filter(total_seq_processed >0) %>%
  arrange(desc(total_seq_processed))
head(loc_variant_df2)
```



# variant, location plot for top 20 countries
```{r}
options(repr.plot.width = 15, repr.plot.height = 30)
loc_variant_df2 %>%
  head(20) %>%
  group_by(location) %>%
  ggplot(mapping = aes(x = total_seq_processed, y = reorder(location, total_seq_processed), fill = total_seq_processed)) + 
  geom_col(position = "dodge", color = "Black") + 
  scale_fill_gradient(low = "#132B43", high = "#56B1F7", space = "Lab", na.value = "gray50", guide = "colourbar", aesthetics = "fill") +
  geom_text(aes(label = total_seq_processed),hjust = -0.1, nudge_x = 0.5, size = 4, angle = 0) +
  labs(title = "COVID Variants vs Total number of Sequences Processed for Top 20 Countries", x = "Total Sequences Processed", y = "Countries", fill = "Total Seq. Processed") +
  scale_x_continuous(labels = scales::comma, limit = c(0, 3000000)) + 
  theme(axis.text = element_text(size = 12)) +
  theme(axis.title = element_text(size = 20), title = element_text(size = 20))
```

# filter Omicron variant
```{r}
omicron_variant_df <- covid_df %>%
  filter(variant == "Omicron")

head(omicron_variant_df)
```


# Omicron data group by location
```{r}
omicron_variant_df <- omicron_variant_df %>%
  group_by(location) %>%
  summarize(total_seq_processed = sum(num_seq_processed)) %>%
  arrange(desc(total_seq_processed))

head(omicron_variant_df)
```

# plot
```{r}
options(repr.plot.width = 15, repr.plot.height = 30)
covid_df %>%
  filter(variant == "Omicron") %>%
  group_by(location) %>%
  summarize(total_seq_processed = sum(num_seq_processed)) %>%
  ggplot(aes(x = total_seq_processed, y = location)) + 
  geom_col(position = "dodge",fill = "Purple" , color = "Black") + 
  labs(title = "Countrywise Omicron COVID Variant vs Total number of Sequences Processed", x = "Total Sequences Processed", y = "Countries", fill = "Countries") +
  scale_x_continuous(labels = scales::comma) + 
  theme(axis.text = element_text(size = 12)) +
  theme(axis.title = element_text(size = 20), title = element_text(size = 20))

```
```{r}
covid_df %>%
  filter(variant == "Omicron") %>%
  group_by(location) %>%
  summarize(total_seq_processed = sum(num_seq_processed)) %>%
  arrange(desc(total_seq_processed)) %>%
  head(10) %>%
  ggplot(aes(x = total_seq_processed, y = location)) + 
  geom_col(position = "dodge",fill = "Purple" , color = "Black") + 
  geom_text(aes(label = total_seq_processed),hjust = -0.1, nudge_x = 0.5, size = 4, angle = 0) +
  labs(title = "Omicron COVID Variant vs Total number of Sequences Processed in Top 10 Countries", x = "Total Sequences Processed", y = "Countries", fill = "Countries") +
  scale_x_continuous(labels = scales::comma, limit = c(0, 80000)) + 
  theme(axis.text = element_text(size = 15)) +
  theme(axis.title = element_text(size = 18), title = element_text(size = 18))
```
